---
title: "Numerical Summary"
author: "Michael Geuenich"
date: "July 28, 2020"
output: html_document
params:
  basel_type: "../../output/phoenix/astir_assignments/basel_astir_assignments.csv"
  basel_metadata: "../../data-raw/metadata/basel_PatientMetadata.csv"

  zurich_type: "../../output/phoenix/astir_assignments/zurich1_astir_assignments.csv"
  zurich_metadata: "../../data-raw/metadata/zurich1_PatientMetadata.csv"

  schapiro_type: "../../output/phoenix/astir_assignments/schapiro_astir_assignments.csv"
  
  wagner_type: "../../output/phoenix/astir_assignments/wagner_astir_assignments.csv"
  wagner_metadata: "../../data-raw/metadata/wagner-clinical-characteristics.csv"
  
  lin_type: "../../output/phoenix/astir_assignments/lin_cycif_astir_assignments.csv"
---

```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(SingleCellExperiment)
library(scater)
library(knitr)
library(DT)
library(devtools)
library(tidyverse)
devtools::load_all("../../../taproom/")
source("../../scripts/functions.R")
```


```{r include=FALSE,message=F}
basel_types <- read_csv(params$basel_type) %>% 
  mutate(cell_type = get_celltypes(select(., -X1))) %>% 
  select(X1, cell_type) %>% 
  mutate(cohort = "basel")

basel_metadata <- read_csv(params$basel_metadata)

schapiro_types <- read_csv(params$schapiro_type) %>% 
  mutate(cell_type = get_celltypes(select(., -X1))) %>% 
  select(X1, cell_type) %>% 
  mutate(cohort = "schapiro")


wagner_types <- read_csv(params$wagner_type) %>% 
  mutate(cell_type = get_celltypes(select(., -X1))) %>% 
  select(X1, cell_type) %>% 
  mutate(cohort = "wagner")


wagner_metadata <- read_csv(params$wagner_metadata)

zurich_types <- read_csv(params$zurich_type) %>% 
  mutate(cell_type = get_celltypes(select(., -X1))) %>% 
  select(X1, cell_type) %>% 
  mutate(cohort = "zurich")


zurich_metadata <- read_csv(params$zurich_metadata)

lin_types <- read_csv(params$lin_type) %>%
  mutate(cell_type = taproom::get_celltypes(select(., -X1), thresh = 0.5)) %>% 
  select(X1, cell_type) %>% 
  mutate(cohort = "lin")

cell_types <- rbind(basel_types, schapiro_types, wagner_types, zurich_types, lin_types)

type_summary <- cell_types %>% 
  group_by(cohort, cell_type) %>% 
  tally() %>% 
  ungroup() %>% group_by(cohort) %>% 
  mutate(total = sum(n)) %>% 
  mutate(Freq = n/total) %>% 
  dplyr::rename(cells_per_type = n, total_per_cohort = total, Frequency = Freq)
```

Total number of cells for: 

- Basel: `r type_summary %>% filter(cohort == "basel") %>% pull(total_per_cohort) %>% unique()`
- Schapiro: `r type_summary %>% filter(cohort == "schapiro") %>% pull(total_per_cohort) %>% unique()`
- Wagner: `r type_summary %>% filter(cohort == "wagner") %>% pull(total_per_cohort) %>% unique()`
- Zurich: `r type_summary %>% filter(cohort == "zurich") %>% pull(total_per_cohort) %>% unique()`
- Lin: `r type_summary %>% filter(cohort == "lin") %>% pull(total_per_cohort) %>% unique()`

Assignment summary
```{r}
datatable(type_summary)
```




### Metadata
Number of basel patients `r length(unique(basel_metadata$PID))`.
Number of basel cores `r length(unique(basel_metadata$core))`.

Number of zurich1 patients `r length(unique(zurich_metadata$PID))`
Number of zurich1 cores `r length(unique(zurich_metadata$core))`

Number of wagner patients `r length(unique(wagner_metadata[["Patient ID"]]))`

Number of schapiro patients `r length(unique(sub("_[^_]+$", "", schapiro_types$X1)))`


### Tumour vs no tumor
```{r}
datatable(as.data.frame(table(basel_metadata$diseasestatus)))
datatable(as.data.frame(table(zurich_metadata$location)))
```

### Patient subtypes
```{r}
datatable(as.data.frame(table(zurich_metadata$Subtype)))
datatable(as.data.frame(table(basel_metadata$clinical_type)))
```


Get number of cell types by tumour vs non-tumour, ER & PR status
```{r warning=F, message=F}
basel_types <- basel_types %>% 
  mutate(core = sub("_[^_]+$", "", X1))

zurich_types <- zurich_types %>% 
  mutate(core = sub("_[^_]+$", "", X1))

basel_covariates <- c("diseasestatus", "ERStatus", "PRStatus", "HER2Status", "clinical_type")
zurich_covariates <- c("location", "ERStatus", "PRStatus", "clinical_type")

basel_freq <- left_join(basel_types, 
                        select(basel_metadata, core, basel_covariates), by = "core")

zurich_freq <- left_join(zurich_types,
                         select(zurich_metadata, core, zurich_covariates), by = "core")


## def function
summarise_covariates <- function(df, grouping){
  cell_type_sum <- df %>% 
    group_by(.data[[grouping]]) %>% 
    count(cell_type)
  
  overall_sum <- cell_type_sum %>% 
    group_by(.data[[grouping]]) %>% 
    summarise(total = sum(n))
  
  cell_type_sum <- dplyr::rename(cell_type_sum, "Number of cells" = n)
  overall_sum <- dplyr::rename(overall_sum, "Number of cells" = total)
  
  list(cell_type_sum, overall_sum)
}
### end 

basel <- lapply(basel_covariates, function(x){summarise_covariates(basel_freq, x)})
zurich <- lapply(zurich_covariates, function(x){summarise_covariates(zurich_freq, x)})
```

### Basel
```{r}
datatable(basel[[1]][[1]])
datatable(basel[[1]][[2]])
datatable(basel[[2]][[1]])
datatable(basel[[2]][[2]])
datatable(basel[[3]][[1]])
datatable(basel[[3]][[2]])
datatable(basel[[4]][[1]])
datatable(basel[[4]][[2]])
```

### Zurich
```{r}
datatable(zurich[[1]][[1]])
datatable(zurich[[1]][[2]])
datatable(zurich[[2]][[1]])
datatable(zurich[[2]][[2]])
datatable(zurich[[3]][[1]])
datatable(zurich[[3]][[2]])
datatable(zurich[[4]][[1]])
datatable(zurich[[4]][[2]])

```